A meta-analysis of fungicide effects on mungbean powdery mildew

### 2021-05-19 04_Severity_MetaAnalysis.Rmd
define logit functions and simplify explainations towards using logit transformation

# Data
PM_dat_D <-
   read.csv("cache/PM_disease_clean_data.csv", stringsAsFactors = FALSE)
```

@@

#logit transformed percentages
PM_dat_D$yi_logit <- log(
   (PM_dat_D$yi_percent) / 
      (1 - (PM_dat_D$yi_percent))
   )
hist(PM_dat_D$yi_logit)
PM_dat_D$vi_logit <-
   (PM_dat_D$vi_percent) /
   (PM_dat_D$yi_percent * (1 - PM_dat_D$yi_percent)) ^ 2

@@

Finally lets assess the logit transformed percent severity
```{r Metafor-Sev-analysis_log, cache=TRUE}
# remove controls for response ratio meta-analysis
PMsev_mv_log <- rma.mv(
   yi = yi_log,
   V = yi_log,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "US",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mv_log)
plot(rstandard(PMsev_mv_log)$z, ylim=c(-3,3), pch=19)
data.frame(cbind((1-exp(PMsev_mv_log$b)),
                              (1-exp(PMsev_mv_log$ci.lb)),
                              (1-exp(PMsev_mv_log$ci.ub))))
```

@@

1- LaplacesDemon::invlogit(PMsev_mvPer$b)
mean(PM_dat_D[PM_dat_D$spray_management == "control","PM_final_severity"])
mean(PM_dat_D[PM_dat_D$spray_management == "Recommended_plus","PM_final_severity"])
```
This produces estimates which are too high and likely incorrect

```{r Metafor-Sev-analysis_logit, cache=TRUE}
#logit transformed 1- 9 values
PM_dat_D$yi_logit <- log(
   ((PM_dat_D$PM_final_severity-1)/8) / 
      (1 - ((PM_dat_D$PM_final_severity-1)/8))
   )

PM_dat_D$vi_logit <-
   (PM_dat_D$vi) /
   (((PM_dat_D$PM_final_severity-1)/8) * (1 -((PM_dat_D$PM_final_severity-1)/8))) ^ 2

# Correct -Inf to zeros
PM_dat_D[PM_dat_D$yi_logit == -Inf |
            PM_dat_D$yi_logit == Inf ,"yi_logit"] <- 0
PM_dat_D[is.na(PM_dat_D$vi_logit == Inf),"vi_logit"] <- 0

PM_dat_D[PM_dat_D$vi_logit <= 0,"vi_logit"] <- 0.001
# remove controls for response ratio meta-analysis
PMsev_mvPer <- rma.mv(
   yi = yi_logit,
   V = yi_logit,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mvPer)
```

```{r}
#logit transformed
PM_dat_D$yi_logit <- log(
   (PM_dat_D$yi_percent) / 
      (1 - (PM_dat_D$yi_percent))
   )

1- LaplacesDemon::invlogit(PMsev_mvPer$b)
mean(PM_dat_D[PM_dat_D$spray_management == "control","PM_final_severity"])
mean(PM_dat_D[PM_dat_D$spray_management == "Recommended_plus","PM_final_severity"])

>>>>

# Data
PM_dat_D <-
   read.csv("cache/PM_disease_clean_data.csv", stringsAsFactors = FALSE)

# functions
logit <- function(x){
   log((x)/(1 - x))
}

inv_logit <- function(x){
   exp(x)/(1 + exp(x))
}

@@

#logit transformed percentages
PM_dat_D$yi_logit <- logit(PM_dat_D$yi_percent)

hist(PM_dat_D$yi_logit)
PM_dat_D$vi_logit <-
   (PM_dat_D$vi_percent) /
   (PM_dat_D$yi_percent * (1 - PM_dat_D$yi_percent)) ^ 2

@@

However square severity is not common when evaluating a response that was obtained on a scale.
Log transformed is a more common method of evaluating non-normal disease severity.

Lets assess the log transformed percent severity
```{r Metafor-Sev-analysis_log, cache=TRUE}
# remove controls for response ratio meta-analysis
PMsev_mv_log <- rma.mv(
   yi = yi_log,
   V = yi_log,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "US",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mv_log)
plot(rstandard(PMsev_mv_log)$z, ylim=c(-3.2,3), pch=19)
data.frame(Estimates = (1-exp(PMsev_mv_log$b)),
           lowerCI = (1-exp(PMsev_mv_log$ci.lb)),
           upperCI = (1-exp(PMsev_mv_log$ci.ub)))
```
The plot of the residuals show the model is not normally distributed.
In addition the estimates don't reflect what can be [observed in the raw data][Disease severity range for each spray schedule].
Log transformed percent severity is therefore no a suitable transformation

@@

# plot the model residuals
plot(rstandard(PMsev_mvPer)$z, ylim=c(-3,3), pch=19)

1 - exp(PMsev_mvPer$b + PMsev_mvPer$b[1])/(1+exp(PMsev_mvPer$b + PMsev_mvPer$b[1]))
logit_estimates <- ifelse(PMsev_mvPer$b == PMsev_mvPer$b[1], PMsev_mvPer$b, PMsev_mvPer$b + PMsev_mvPer$b[1])
1 - inv_logit(logit_estimates)
```
